A comparative study of batch construction strategies \\ for recurrent neural networks in MXNet

### 2017-04-19 paper/paper.tex
semi-final, 3 WERs missing

    A straight-forward strategy to minimize zero padding is to sort the sequences by length and to partition them into
    batches afterwards. However, there are significant drawbacks to this method. First, the sequence order remains constant in each epoch 
    and therefore the intra-batch variability is very low since the same sequences are usually combined into the same batch. Second, the strategy favors putting very similar sequences into the same batch, since short sequences often tend to share other properties.
    One way to overcome this limitation was proposed within TensorFlow and is also used as recommended 
    behavior in MXNet The idea is to perform a \textit{bucketing} of the training corpus, where each bucket represents
    a range of sequence lengths and each training sample is then assigned to the bucket that corresponds to its length.
    Afterwards a batch is constructed from a randomly chosen bucket by selecting all or some sequences from it.
    The concept somehow mitigates the issue of zero padding if suitable length ranges can be defined, while still allowing for 
    some level of randomness at least when sequences are selected within a bucket. However, buckets have to be made very large 
    in order to ensure a sufficiently large variability within batches. On the other hand, making buckets too large will 
    increase training time due to irrelevant computations on zero padded frames. Setting these hyper-parameters correctly is
    therefore of fundamental importance for fast and robust acoustic model training.

@@

  We extracted 16-dimensional MFCC vectors as in \cite{menne16:chime4System} from the six sub-corpora
  and used them as features for the neural network models. The allophones of the lexicon were clustered into 
  1500 state labels using a classification and regression tree (CART).
  We trained a three state HMM without skip transitions applied it in a forced alignment mode to the training data in order to obtain a frame-wise labeling. A three layer 
  bi-directional LSTM followed a softmax layer was trained to minimize the cross-entropy of the softmax output and the CART label sequence at each frame. Optimization was done with the Adam optimizer and a constant learning rate of 0.01.

@@

  \section{Conclusions}
    In this work we presented a novel strategy to construct sequence-level batches for recurrent 
    neural network training. Most frameworks rely on a bucketing approach by clustering sequences of similar length into bins and to create batches from each bin individually. We showed that we can achieve a better runtime performance using a simpler batch design, by partitioning a shuffled sequence order and to sort the partitions in alternating order. The method was evaluated on the ChiME-4 noisy speech recognition task and 
    compared to the bucketing approach of MXNet.

>>>>

    A straight-forward strategy to minimize zero padding is to sort the sequences by length and to partition them into
    batches afterwards. However, there are significant drawbacks to this method. First, the sequence order remains constant in each epoch 
    and therefore the intra-batch variability is very low since the same sequences are usually combined into the same batch. Second, the strategy favors putting very similar sequences into the same batch, since short sequences often tend to share other properties.
    One way to overcome this limitation was proposed within TensorFlow and is also used as recommended 
    behavior in MXNet. The idea is to perform a \textit{bucketing} of the training corpus, where each bucket represents
    a range of sequence lengths and each training sample is then assigned to the bucket that corresponds to its length.
    Afterwards a batch is constructed from a randomly chosen bucket by selecting all or some sequences from it.
    The concept somehow mitigates the issue of zero padding if suitable length ranges can be defined, while still allowing for 
    some level of randomness at least when sequences are selected within a bucket. However, buckets have to be made very large 
    in order to ensure a sufficiently large variability within batches. On the other hand, making buckets too large will 
    increase training time due to irrelevant computations on zero padded frames. Setting these hyper-parameters correctly is
    therefore of fundamental importance for fast and robust acoustic model training.

@@

  We extracted 16-dimensional MFCC vectors as in \cite{menne16:chime4System} from the six sub-corpora
  and used them as features for the neural network models. The allophones of the lexicon were clustered into 
  1500 state labels using a classification and regression tree (CART).
  We trained a three state HMM without skip transitions applied it in a forced alignment mode to the training data in order to obtain a frame-wise labeling. A three layer 
  bi-directional LSTM followed a softmax layer was trained to minimize the cross-entropy of the softmax output and the CART label sequence at each frame. Optimization was done with the Adam optimizer and a constant learning rate of 0.01. We used MXNet \cite{mxnet} for experiments using the bucketing approach and RETURNN \cite{doetsch2017:returnn} for the other methods.

@@

  \section{Conclusions}
    In this work we presented a novel strategy to construct sequence-level batches for recurrent 
    neural network acoustic model training. While not much attention is given to the topic 
    of batch construction, we demonstrate that different strategies can lead to large variations 
    both in training time and recognition performance. Most deep-learning frameworks rely on a bucketing approach by clustering sequences of similar length into bins and to create batches from each bin individually. We showed that we can achieve a better runtime performance using a simpler batch design, by partitioning a shuffled sequence order and to sort the partitions in alternating order. The method was evaluated on the ChiME-4 noisy speech recognition task and 
    compared to standard approaches like random sequence shuffling and the bucketing approach of MXNet,
    where our method was able to reach a better trade-off between training time and recognition performance while being easier to parametrize than the bucketing method.
